---
title: " Statistics Homework"
author: "Niall Horan"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Lesson 1
###1.13
Find the mean and median of 123,116,122,110,175,126,125,111,118,117.
$$\bar{x}=\frac{123+116+\dots +117}{10}=124.3$$
$$\tilde x=110,111,116,117,\textbf{118,122},123,125,126,175=120$$

```{r}
x<-c(110,111,116,117,118,122,123,125,126,175)
mean(x)
median(x)
```

Notice the difference between the two.  The mean is larger because the value of 175.

###1.14
data set is 572,572,573,568,569,575,565,570.

####a. Find the mean and median.
$$\bar{x}=\frac{572+572+\dots +570}{8}=570.5$$
$$\tilde x=565,568,569,\textbf{570,572},572,573,575=571$$

####b. Find the sample variance, standard deviation, and range.
$$s^2=\sum_{i=1}^n\frac{(x_i-\bar{x})^2}{n-1}=\frac{1}{7}[(565-570.5)^2+(568-570.5)^2+\dots +(575-570.5)^2]=10$$
$$s=\sqrt{10}=3.16$$


```{r}
x<-c(572,572,573,568,569,575,565,570)
mean(x)
median(x)
var(x)
sd(x)
range(x)
```

####c. Although the average is slightly greater than 570, the standard deviation shows that a good amount of tires will be substantailly less than 570 or greater than 570.  

##Lesson 2
###1.18
a. Stem and Leaf
b. histogram
c. mean, median and standard deviation.
```{r , echo=FALSE, fig.width=7,fig.height=6,fig.cap="Stem and Leaf"}
x<-c(23,	60,	79,	32,	57,	74,	52,	70,	82,	36,	80,	77,	81,	95,	41,	65,	92,	85,	55,	76,	52,	10,	64,	75,	78,	25,	80,	98,	81,	67,	41,	71,	83,	54,	64,	72,	88,	62,	74,	43,	60,	78,	89,	76,	84,	48,	84,	90,	15,	79,	34,	67,	17,	82,	69,	74,	63,	80,	85,	61)
stem(x)
```

```{r , echo=FALSE, fig.width=7,fig.height=6,fig.cap="Histogram and Boxplot"}
par(mfrow=c(1,2))
hist(x)
boxplot(x, ylab="grades")
par(mfrow=c(1,1))
stats<-data.frame("mean"=mean(x), "median"=median(x),"standard deviation"=sd(x))
stats
```

###1.27

a. 
```{r}
x<-c(700,1000,1300)
y<-c(210,325,375)
plot(x,y, xlab="Load Level", ylab="sample Means")
y1<-c(145,250,150)
y2<-c(105,195,180)
y3<-c(260,375,420)
y4<-c(330,480,750)
df<-data.frame("Level"=x,"y1"=y1,"y2"=y2,"y3"=y3,"y4"=y4)
library(ggplot2)
library(tidyr)
library(dplyr)
newDF<-gather(data=df,'y1','y2','y3','y4', key='specimen', value='weight')
ggplot(data=newDF,aes(x=Level,y=weight,color=specimen))+geom_line()
```

##Lesson 3.
###8.17

$$n=16 \hspace{1 cm} \mu=50 \hspace{1 cm}  \sigma=5$$
What is the probability $$\mu_{\bar{x}}-1.9\sigma_\bar{x} <\bar{x}<\mu_{\bar{x}}-0.4\sigma_\bar{x}$$
There are two ways to do this.  The easy way is to understand that this is the same as saying 
$p(-1.9<z<-0.4)$ and using the standard normal we find this to be 0.315.  

The other way to do this is the long way using the values given.

$$\mu_{\bar{x}}-1.9\sigma_\bar{x}=50-5*(1.9)/\sqrt{16} \hspace{.25cm} \text{and}\hspace{.25cm} \mu_{\bar{x}}-0.4\sigma_\bar{x}=50-5*(0.4)/\sqrt{16}$$
so now we just want to know the probablity $47.625<\bar{x}<49.5$
Using $z=\frac{\bar{x}-\mu}{\sigma/\sqrt{n}}$
```{r}
pnorm(-.4,0,1)-pnorm(-1.9,0,1)#easy
pnorm(49.5,50,5/4)-pnorm(47.625,50,5/4)#hard
```

####8.19
$\mu=78.3 \hspace{.5cm} \sigma=5.6$
a. Sample Variance = $$\frac{\sigma}{\sqrt{n}}$$
$n$ changes from 64 to 96. $\frac{5.6}{\sqrt{64}}=0.7 \hspace{.5cm} \frac{5.6}{\sqrt{196}}=0.4$ SO the variance is reduced from 0.49 to .16.
####8.22
$$\mu=174.5 \hspace{.5cm} \sigma=6.9 \hspace{.5cm} n=25$$
a. $\mu_{\bar{x}}=174.5 \hspace{.5cm} \sigma_{\bar{x}}=\frac{6.9}{\sqrt{25}}=0.276$
b. Find the proportion or probability the sample mean falls between 172.5 and 175.8.
$$P(172.5<\bar{x}<175.8)=P(Z<\frac{175.8-174.5}{\frac{6.9}{\sqrt{25}}})-P(Z<\frac{172.5-174.5}{\frac{6.9}{\sqrt{25}}})$$
$$=P(z<0.942029)-P(z<-1.449275)=0.8269111-0.0736304=0.75$$
c. Just like b. but $p(\bar{x}<172)=0.035$
```{r}
pnorm(175.8,174.5,6.9/5)-pnorm(172.5,174.5,6.9/5)#b
pnorm(172,174.5,6.9/5)#c
```

###8.25
a.
$$\mu=7 years, \hspace{.5cm}\sigma=1 year \hspace{.5cm} n=9\hspace{.5cm} p(6.4<\bar{x}<7.2)$$
$$P(6.4<\bar{x}<7.2)=P(Z<\frac{7.2-7}{\frac{1}{\sqrt{9}}})-P(Z<\frac{6.4-7}{\frac{1}{\sqrt{9}}})=.69$$

b.$P(\bar{X}>\bar{x})=0.15 = 1-P(Z<\frac{\bar{x}-7}{1/\sqrt{9}}) \hspace{.5cm} Z(.85)=1.036$ so  $\bar{x}=7.35$
```{r}
qnorm(.85,0,1)*(1/3)+7
```
###8.29

```{r}
mu<-c(72,28)
sigma<-c(10,5)
n<-c(64,100)
df<-data.frame(mu,sigma,n)
dnames<-c('terrier','poodle')
row.names(df)<-dnames
df
(72-28-44.2)/sqrt(100/68+25/100)
```
$$Z=\frac{44.2-(72-28)}{\sqrt{(100/64)+(25/100)}}=-0.15\hspace{1cm}p(z<0.15)=.55$$
##Lesson 4
###8.41

$n=25 \hspace{.5cm} \sigma^2=6 \hspace{.5cm}\text{Normal Population}$

$P(S^2>9.1)$

$\chi^2=\frac{(n-1)s^2}{\sigma^2}=\frac{(25-1)9.1}{6}=36.4$

$P(\chi^2>36.4)=.05 \hspace{.25cm} \nu=24 \text{  from tables}$

$P(3.462<s^2<10.745)$

$p(\frac{(25-1)3.462}{6}<\chi^2<\frac{(25-1)10.745}{6})$

$P(13.85<\chi^2<42.98)\hspace{.25cm}\nu=24$

$.99-.05=.94$


```{r}
plot(1:60,dchisq(1:60,24), type="l",xlab=expression(chi^2), ylab=expression(f(x)))
#a
1-pchisq(36.4,24)     

#b.
pchisq(42.98,24)-pchisq(13.85,24)
```

###8.42

$$\chi^2=\frac{(n-1)s^2}{\sigma^2}=\frac{(20-1)20}{8}=47.5$$

$p(\chi^2>47.5)\hspace{.25cm}{\nu=19}\hspace{.25cm}=0.0003$

S it is very unlikely for this to occur!
```{r}
19*20/8
 1-pchisq(47.5,19)
```

###8.49

$\mu=20$ and the sample standard deviation of the mean is unknown se we approximate it with s. $\frac{s}{\sqrt(n)}=\frac{4.1}{\sqrt{9}}=1.4$, so if the standard deviation of the sample mean with $n=9$ is 1.4, to get a mean of 24 is very unlikely(well over two standard deviations aways.)

###8.50
Claim is $\mu=0.5$
$$t=\frac{\bar{x}-\mu}{s/\sqrt{n}}$$ Where the sample mean is 0.475, $S$=0.183, and n=8.
t=-.049. $p(t<-.049)=0.48 $ ( with 7 degrees of freedom)
So pretty likely to occur.

```{r}
x<-c(.6,.7,.7,.3,.4,.5,.4,.2)
mean(x)
sd(x)
(.475-.5)/.18/sqrt(8)
pt(-.05,df=7)
```

###8.53
Two Mines with the below data. Can we conlude the population variances are equal?
$s^2_1=15750$
$s^2_2=10920$
Don't really look the same.  But we should use the f-test!!!
$$F=\frac{s_1^2}{s_2^2}$$ with $\nu_1=n_1-1$ and $\nu_2=n_2-1$ degrees of freedom.
Lets keep it right tailed and put the larger value on the top.

$F=\frac{15750}{10920}=1.44$
with $\nu=5 \text{ and } 4$ our tables say that 5% of values are greater than 4, so this looks pretty legit.

```{r}
m1<-c(8260,8130,8350,8070,8340)
m2<-c(7950,7890,7900,8140,7920,7840)
var(m1)
var(m2)
```

##Lesson 5

###9.2

Need to find the 96% confidence interval.

$\sigma = 40 \hspace{.5cm} n=30 \hspace{.5cm} \bar{x}=780 \hspace{.5cm} \alpha=.04 \hspace{.5cm} z_{\alpha/2}=2.05$

$\bar{x}\pm z_{\alpha/2}\times\sigma/\sqrt{n}$

$780\pm 2.0540\times 40/\sqrt{30}$

$(765.0289 \hspace{.5cm} 794.9711)$

###9.4
98% confidence interval.

$n=50, \thinspace  \bar{x}=174.5, \thinspace  \sigma=6.9, \thinspace \alpha=.02$

$\bar{x}\pm z_{\alpha/2}\times\sigma/\sqrt{n}$

$174.5\pm 2.34\times6.9/\sqrt{50}$

$(172.2166, \thinspace 176.7834)$

###9.7
Solve for n. Error=.0005, $\alpha=.05$
Use thereom 9.2.
$n=(\frac{z_{\alpha/2}\times \sigma}{e})^2$
34.6 or 35

###9.9
$n<30$ and we do not know populatoin standard deviaton so we are going to use student $t$.\n
$\bar{x}=11.3, \thinspace s=2.45, \thinspace n=20, \thinspace \alpha=.05$ 
$$\bar{x}\pm t_{\alpha/2}\times s/\sqrt{n}$$
$$\bar{11.3}\pm 2.093\times2.45/\sqrt{20}$$
( 10.15338, 12.44662)

###9.13

$n<30$ and we do not know populatoin standard deviaton so we are going to use student $t$.\n
$\bar{x}=48.50, \thinspace s=1.5, \thinspace n=12, \thinspace \alpha=.10$ 
$$\bar{x}\pm t_{\alpha/2}\times s/\sqrt{n}$$

$$48.5\pm 1.80\times 1.5/\sqrt{12}$$
( 47.72236 49.27764)

###9.17
95% prediction interval from 9.9.\n
$n<30$ and we do not know populatoin standard deviaton so we are going to use student $t$.\n
$\bar{x}=11.3, \thinspace s=2.45, \thinspace n=20, \thinspace \alpha=.05$
$$\bar{x}\pm t_{\alpha/2}\times s\sqrt{1+1/n}$$
$$11.3\pm 2.09\times 2.45\sqrt{1+1/20}$$
(6.053049 16.546951)
